Foundations for Inference

### 2018-10-21 DATA606/Foundations for Inference.Rmd
Minor change
More work on homework

(b) Suppose you read online that children first count to 10 successfully when they are 32 months old, on average. Perform a hypothesis test to evaluate if these data provide convincing evidence that the average age at which gifted children first count to 10 successfully is less than the general average of 32 months. Use a significance level of 0.10.

(c) Interpret the p-value in context of the hypothesis test and the data.

(d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.

(e) Do your results from the hypothesis test and the confidence interval agree? Explain.

@@

@@

(c) Do your results from the hypothesis test and the confidence interval agree? Explain.

@@

(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Centered around 9,000 hours.**

(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours? **It would be roughly 0.067**

(d) Sketch the two distributions (population and sampling) on the same scale.

(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Yes, but not using the above method.**

>>>>

(b) Suppose you read online that children first count to 10 successfully when they are 32 months old, on average. Perform a hypothesis test to evaluate if these data provide convincing evidence that the average age at which gifted children first count to 10 successfully is less than the general average of 32 months. Use a significance level of 0.10. **$H_0$ is that x_bar >= 32 and $H_a$ is x_bar < 32**

```{r}
confidence_level <- 0.10
n <- 36
x_bar <- 30.69
sigma <- 4.31
mu <- 32

se_x_bar <- sigma / sqrt(n)

z <- (x_bar - mu) / se_x_bar

z_score_for_confidence_level <- 1.645

p_value <- pnorm(z, mean=0, sd=1) * 2

if(z < z_score_for_confidence_level){
  message(paste(round(z, 3), "is less than", z_score_for_confidence_level, "so reject the null hypothesis (p-value:", round(p_value, 3), ")"))
} else {
  message(paste(round(z, 3), "is not less than", z_score_for_confidence_level, "so do not reject the null hypothesis (p-value:", round(p_value, 3), ")"))
}
```

(c) Interpret the p-value in context of the hypothesis test and the data.  **There is a 0.068 probability that the difference between these two means was due to chance.**

(d) Calculate a 90% confidence interval for the average age at which gifted children first count to 10 successfully.
```{r}
# Same method as http://onlinestatbook.com/2/calculators/inverse_normal_dist.html
zinv <- function(p) {
	if (p >= 1 || p <= 0) {
		return(FALSE)
	}
	if (p < .5) {
		t <- p
		theSign <- -1
	}
	else {
		t <- 1 - p
		theSign <- 1
	}

	v <- sqrt(-2.0 * log(t))
	x <- 2.515517 + (v * (0.802853 + v * 0.010328))
	y <- 1 + (v * (1.432788 + v * (0.189269  + v * 0.001308)))
	Q <- theSign * (v - (x / y))
	return(Q)
}

p <- 1 - confidence_level
x1 <- x_bar + sigma * zinv(p)
p2 <- p / 2
x1 <- zinv(.5 - p2)

error <- sigma * -x1
ci_low <- x_bar - error
ci_high <- x_bar + error
```
**Confidence interval: (`r ci_low`, `r ci_high`) or `r x_bar` $\pm$ `r error`**

(e) Do your results from the hypothesis test and the confidence interval agree? Explain. **No.  Since 32 is inside of the confidence interval you would not reject the null hypotheses.  It is the opposite conclusion you reach with the hypothesis test, which indicates to me that I did something wrong.**

@@

```{r}
confidence_level <- 0.10
n <- 36
x_bar <- 118.2
sigma <- 6.5
mu <- 100

se_x_bar <- sigma / sqrt(n)

z <- (x_bar - mu) / se_x_bar

z_score_for_confidence_level <- 1.645

p_value <- pnorm(z, mean=0, sd=1)

if(z < z_score_for_confidence_level){
  message(paste(round(z, 3), "is less than", z_score_for_confidence_level, "so reject the null hypothesis (p-value:", round(p_value, 3), ")"))
} else {
  message(paste(round(z, 3), "is not less than", z_score_for_confidence_level, "so do not reject the null hypothesis (p-value:", round(p_value, 3), ")"))
}
```

@@

```{r}
p <- 1 - confidence_level
x1 <- x_bar + sigma * zinv(p)
p2 <- p / 2
x1 <- zinv(.5 - p2)

error <- sigma * -x1
ci_low <- x_bar - error
ci_high <- x_bar + error
```

**Confidence interval: (`r ci_low`, `r ci_high`) or `r x_bar` $\pm$ `r error`**

(c) Do your results from the hypothesis test and the confidence interval agree? Explain. **No. 100 lies outside the CI which would suggest that we should reject the null hypothesis, but the hypothesis test disagrees.  Again it would appear there is something wrong with my understanding as both should lead you to the same conclusion.**

@@

(b) Describe the distribution of the mean lifespan of 15 light bulbs. **Normal and centered around 9,000 hours.**

(c) What is the probability that the mean lifespan of 15 randomly chosen light bulbs is more than 10,500 hours?

```{r}
n <- 15
se <- sd / sqrt(n)
1 - pnorm(x, mu, se)
```

(d) Sketch the two distributions (population and sampling) on the same scale.
```{r, echo=FALSE}
library(ggplot2)
set.seed(4)
pop <- rnorm(1000000, mu, sd)
sample_pop <- sample(pop, 15)
data <-  rbind(data.frame(x=pop, Distribution=c('Population')), data.frame(x=sample_pop, Distribution=c('Sampling')))

ggplot(data, aes(x=x,color=Distribution)) + geom_density() + theme(axis.title.x = element_blank())
```

(e) Could you estimate the probabilities from parts (a) and (c) if the lifespans of light bulbs had a skewed distribution? **Not using the above method.**
